Open Source, Collaborative Curriculum

### 2017-04-29 siip.tex
lots of edits toward final version

          \begin{document}
          \title{Open Source, Collaborative Curriculum}
          \author{PI: Kathryn Huff\\Co-PI: Neal Davis}
          \maketitle

@@

          Attempts at this have been made before. Most notable among 
          these is Software Carpentry (with which the PI and Co-PI are quite involved). 
          Software Carpentry teaches computational skills for scientists using 
          a shared curriculum hosted on GitHub. Unfortunately, our experience showed that 
          forks of the original material tend to diverge without any strong 
          incentive for contribution back to the master material 
          \cite{wilson_software_2014,wilson_software_2014-1}. We expect that fine-grained modularity 
          of lesson components and a clear dependency graph of prerequisite 
          modules for each lesson may assist in overcoming the challenges 
          encountered by Software Carpentry and others.

          \section{Proposed Solution}
          We propose a small-scale proof-of-concept for collaborative, open 
          source, curriculum development to improve the transfer of lessons 
          learned between instructors of the same course (either at a single 
          university or among different campuses). This prototype collaboration 
          will provide a template which could be adopted for collaboration 
          among faculty teaching courses with an inherently larger scale (e.g. 
          CS101)

          Faculty in Nuclear Engineering from four institutions\footnote{Paul 
          Wilson at (UW Madison), Steven Skutnik (UT Knoxville), Jeremy Roberts 
          (Kansas State University), Anthony Scopatz (University of South 
          Carolina), and Robert Borrelli (University of Idaho)} have agreed to be participants in 
          this prototyping effort. We all currently use 
          \href{https://github.com}{GitHub} to store, 
          revise, and collaborate on research, especially source code. 
          Additionally, this select group already started to host our course 
          curricula online as well, but these are typically single author 
          repositories (e.g. \cite{huff_npre412_2017}).

          The participants will collaborate on a master set of learning 
          modules for an upper-division course in nuclear engineering : 
          The Nuclear Fuel Cycle. In the near term, we will develop fine 
          grained lessons which may be mixed and matched to meet the 
          learning objectives. The curriculum will be hosted on GitHub, 
          tested by all of us, and improved continually.

@@

          Education is an inherently distributed system, but this need not be a hinderance to collaboration.

@@

          \subsection{Kick-Off Workshop}
          A two-day kick-off workshop (\$15K) at UIUC will allow the 
          participants to sketch the initial framework of 
          the collaboration and coordinate logistics. The workshop will 
          deliver:

          \begin{itemize} 
                  \item learning objectives\cite{bloom_bloom_1984} for the 
                          course
                  \item a concept map\cite{novak_concept_1990} for the course
                  \item a list of fine-grained lesson topics
                  \item a directed acyclic graph describing dependencies among 
                          lessons
                  \item individual assignments for curriculum development
                  \item decisions to adopt specific raw formats for lesson content
                        (markdown vs. \LaTeX, Jupyter vs MATLAB, etc.)
                  \item a repository structure for organizing the lesson 
                          material (at this stage, compatibility with 
                          frameworks like RELATE 
                          \cite{kloeckner_relate_2017,kloeckner_relate_2017} 
                          and tools like PrarieLearn 
                          \cite{west_prairielearn:_2015} will be considered)
          \end{itemize}

          Invited participants in the first workshop will include:
          \begin{itemize}
                  \item UIUC PI Huff and UIUC Co-PI Davis
                  \item external collaborating Professors Wilson, Roberts, 
                          Borrelli, Skutnik, and Scopatz
                  \item up to 5 UIUC professors who may be likely to extend 
                          this work in the future for their own courses.
          \end{itemize}

@@

          \section{Timeline and Deliverables}
          \subsection{Timeline}
          The timeline will be bookended by two workshops, a kickoff workshop 
          and a retrospective workshop. In the intervening time, consistent 
          communication through GitHub's collaboration framework, a Slack 
          channel, an email listhost, a shared Google Drive folder, and

          \begin{itemize}
                  \item[July 2017] 1st Workshop
                  \item[.] 
                  \item[.] 
                  \item[.] 
                  \item[July 2018] Retrospective Workshop
          \end{itemize}

@@

          \subsubsection{Learning Objectives List}
          \subsubsection{Concept Map}
          \subsubsection{Learning Dependency Graph}
\begin{figure}[htb!]
        \begin{center}
                \includegraphics[height=0.5\textheight]{rosalind.png}
        \end{center}
        \caption{Rosalind.info has a graph of exercises and their dependencies.}
        \label{fig:rosalind}
\end{figure}
          \subsubsection{Lesson Materials}

          \section{Budget}
          The two workshops will be hosted at the Illini Union, where a meeting 
          room will be reserved. Visiting 
          invited workshop participants will be flown to Urbana-Champaign and 
          lodged at the Illini Union Hotel.

\begin{table}[htb]
        \begin{tabularx}{\textwidth}{| X |c|c|c|c|c|}
        \hline
        \textbf{Event} & \textbf{Item} & \textbf{Cost} & \textbf{Units} & \textbf{\$} & \textbf{Notes}\\ 
        \hline
Kick-off Workshop&Meeting Room&744&2&1488&Allerton, all day package\\
&Lodging&150&24&3600&Allerton guest rooms\\
&Flights&800&6&4800&Estimated\\
&Taxi&20&12&240&\\
&Incidentals&100&6&600&\\
&Continental breakfast&8.5&24&204&Working breakfasts, Allerton\\
&Dinner&30&24&720&In Champaign-Urbana\\
\hline
Retrospective Workshop&Meeting Room&50&2&100&Illini Union\\
&Projector&300&4&1200&Illini Union\\
&Lodging&150&24&3600&Illini Union Hotel\\
&Flights&800&6&4800&Estimated\\
&Taxi&20&12&240&\\
&Incidentals&100&6&600&\\
&Continental breakfast&8.5&24&204&University Catering\\
&Coffee Service&32.5&4&130&\\
&Lunch&15&24&360&University Catering\\
&Dinner&30&24&720&In Champaign-Urbana\\
        \hline
        &\textbf{Total}&&&\textbf{23606}&\\
        \hline
\end{tabularx}
\end{table}

>>>>

          \begin{document}
          \title{Open Source, Collaborative Curriculum}
          \author{\textbf{PI: Kathryn Huff}\\%
%          \textbf{Co-PI: Neal Davis}\\%
                  University of Illinois Urbana-Champaign
                  \and
\textbf{Co-PI: Neal Davis}\\University of Illinois Urbana-Champaign
\and
Paul Wilson\\University of Wisconsin - Madison 
\and
          Steven Skutnik\\University of Tennessee - Knoxville
\and
          Anthony Scopatz\\University of South Carolina 
\and
          Jeremy Roberts\\Kansas State University 
\and
          Robert Borrelli\\University of Idaho at Idaho Falls
}
          \maketitle

@@

          Previous attempts at collaborative curriculum development have met with 
          challenges. Most notable among these is Software Carpentry (with 
          which the PI and Co-PI are quite involved) 
          \cite{wilson_software_2014}.  Software Carpentry teaches 
          computational skills for scientists using a shared curriculum hosted 
          on GitHub. Instructors can use and remix the curriculum, but our 
          experience showed that instructors improved on the original material, 
          but lacked a  strong incentive for contribution back to the master 
          material. Thus, versions of the original material tended to diverge 
          rather than converge 
          \cite{wilson_software_2014,wilson_software_2014-1}.

          We expect that 
          fine-grained modularity of lesson components and a clear dependency 
          graph of prerequisite modules for each lesson may assist in 
          overcoming the challenges encountered by Software Carpentry and 
          others.

          \section{Proposed Solution}
          We propose a small-scale proof-of-concept for collaborative, open 
          source, curriculum development to improve the transfer of lessons 
          learned between instructors of the same course (either at a single 
          university or among different campuses). This prototype collaboration 
          will provide a template which could be adopted for collaboration 
          among faculty teaching courses with an inherently larger scale (e.g. 
          CS101).

          Faculty in Nuclear Engineering from five institutions\footnote{
          Paul Wilson at (UW Madison), 
          Steven Skutnik (UT Knoxville), 
          Anthony Scopatz (University of South Carolina), 
          Jeremy Roberts (Kansas State University), 
          and Robert Borrelli (University of Idaho)
          } have agreed to be participants in 
          this prototyping effort. We all currently use 
          \href{https://github.com}{GitHub} to store, 
          revise, and collaborate on research, especially source code. 
          Additionally, this select group already started to host our course 
          curricula online as well, but these are typically single author 
          repositories (e.g. \cite{huff_npre412_2017}).

          The participants will collaborate on a master set of learning 
          modules for an upper-division course in nuclear engineering : 
          on The Nuclear Fuel Cycle. In the near term, we will develop 
          fine-grained lessons and assessments which may be mixed and matched 
          to meet the learning objectives. The curriculum will be hosted on 
          GitHub, tested by all of us, and improved continually.

          Prof. Neal Davis, Co-PI of this proposal, has contemplated a 
          Collaborative Computational Curriculum for some time 
          \cite{davis_ui-ccc_2016}. Since he 
          teaches large-scale 
          computational curriculum at UIUC 
          he is an ideal person to potentially scale-up this work in the 
          future. Additionally, he has the nuclear engineering background to 
          understand the context of this prototyping effort. Prof. Davis' role 
          in this effort will be to conduct \emph{action research}, observing 
          our process and teasing out potential avenues for future 
          extensions and applications of the process.

@@

          Education is an inherently distributed system, but this need not be a hindrance to collaboration.

@@

          \subsection{Kick-Off Workshop}
          A two-day kick-off workshop hosted at the University of Illinois at 
          Urbana-Champaign will allow the 
          participants to sketch the initial framework of 
          the collaboration and coordinate logistics. Through brainstorming and 
          discussion activities, the workshop will deliver a number of 
          products. These products will be published in an online repository 
          and associated website, both hosted by GitHub.

          Invited participants in the first workshop will include UIUC PI Huff 
          and UIUC Co-PI Davis, excternal collaborating professors Wilson, 
          Skutnik, Roberts, Scopatz, and Borrelli, and up to 2 UIUC professors 
          who may be likely to extend this work in the future for their own 
          courses.

          As indicated in Table \ref{tab:kickoff}, products to be developed in the first workshop include learning 
          objectives\cite{bloom_bloom_1984}, a concept 
          map\cite{novak_concept_1990}, a list of fine-grained lesson topics, a 
          directed acyclic graph describing dependencies among lessons. Additionally, concrete logistics 
          for completion will be established, including individual curriculum 
          development assignments, decisions to adopt specific raw formats 
          for lesson content (markdown vs. \LaTeX, Jupyter vs MATLAB, etc.), 
          and a repository structure for organizing the lesson material. At 
          this workshop, interest and compatibility with frameworks like RELATE 
          \cite{kloeckner_relate_2017,kloeckner_relate_2017} and tools like 
          PrarieLearn \cite{west_prairielearn:_2015} will be considered.

          \begin{table}[h!]
                  \centering
          \begin{tabularx}{\textwidth}{r|l|l|X}
                                  \multicolumn{4}{c}{\textbf{Day 1}}\\
                                  \hline
                          \textbf{Time}& \textbf{Activity}& \textbf{Lead}& \textbf{Deliverable}\\
                                  \hline
8:00& Welcome and Breakfast& Huff \& Davis& \\
9:00& Vision Lightning Talk& Wilson& Idea Collection\\
9:15& Vision Lightning Talk& Skutnik& \\
9:30& Vision Lightning Talk& Scopatz& \\
9:45& Vision Lightning Talk& Roberts& \\
10:00& Vision Lightning Talk& Borrelli& \\
10:15& Vision Lightning Talk& Huff& \\
10:30& Break&&  \\
10:45& Concept Mapping Exercise& All& Curriculum Concept Map\\
12:00& Lunch& All& \\
13:00& Brainstorm Learning Objectives& All& Learning Objectives List\\
15:00& Break&&  \\
15:15& Refine Lesson Topics& All& Lesson Topics List\\
16:00& Discussion of Formats for Content& All&  \\
17:00& Break&&  \\
19:00& Dinner&&  \\
\hline
                                  \multicolumn{4}{c}{}\\
                                  \multicolumn{4}{c}{\textbf{Day 2}}\\
                                  \hline
                          \textbf{Time}& \textbf{Activity}& \textbf{Lead}& \textbf{Deliverable}\\
                                  \hline
8:00& Review and Breakfast& Huff \& Davis& \\
9:00& Lesson Dependency Exercise& All& Lesson Dependency Graph\\
10:30& Break&&  \\
10:45& Lesson Dependency Exercise& All& \\
12:00& Lunch& All& \\
13:00& Repository Initialization&&  Template Workspace\\
14:00& Determination of Milestones& All& Work Plan\\
15:00& Break&&  \\
15:15& Assigment of Responsibilities& All&  \\
16:00& Review and Wrap-up& All&  \\
17:00& Break&&  \\
19:00& Dinner&&  \\
\hline
                  \end{tabularx}
                  \caption{Preliminary schedule for the Kick-off Workshop}
                  \label{tab:kickoff}
          \end{table}

@@

          \section{Timeline and Deliverables}
          \subsection{Timeline}
          The timeline will be bookended by two workshops, a kickoff workshop 
          and a retrospective workshop. In the intervening time, consistent 
          communication through GitHub's collaboration framework, a Slack 
          channel, an email listhost, a shared Google Drive folder, and monthly 
          Google Hangouts will drive this work.

          \begin{itemize}
                  \item[Jul 2017] 1st Workshop
                  \item[Aug 2017 - Jun 2018] Monthly Video Conferences 
                  \item[Jul 2018] Retrospective Workshop
          \end{itemize}

@@

          \subsubsection{Learning Objectives List}
          \subsubsection{Concept Map}
          \subsubsection{Learning Dependency Graph}
\begin{figure}[ht!]
        \begin{center}
                \includegraphics[height=0.5\textheight]{rosalind.png}
        \end{center}
        \caption{Rosalind.info, an educational bioinformatics resource, 
        organizes exercises based on their dependencies.}
        \label{fig:rosalind}
\end{figure}
          \subsubsection{Lesson Materials}

          \section{Budget}
          The first workshop will be hosted at the Allerton Retreat Center, 
          where a small meeting room will be reserved. Visiting 
          invited workshop participants will be flown to Urbana-Champaign and 
          lodged at the Illini Union Hotel.

\begin{table}[h!]
        \begin{tabularx}{\textwidth}{| X |c|c|c|c|c|}
        \hline
        \textbf{Event} & \textbf{Item} & \textbf{Cost} & \textbf{Units} & \textbf{\$} & \textbf{Notes}\\ 
        \hline
Kick-off Workshop&Meeting Room&744&2&1488&Allerton, all day package\\
&Lodging&150&24&3600&Allerton guest rooms\\
&Flights&800&6&4800&Estimated\\
&Taxi&20&12&240&\\
&Incidentals&100&6&600&\\
&Continental breakfast&8.5&24&204&Working breakfasts, Allerton\\
&Dinner&30&24&720&In Champaign-Urbana\\
\hline
Retrospective Workshop&Meeting Room&50&2&100&Illini Union\\
&Projector&300&4&1200&Illini Union\\
&Lodging&150&24&3600&Illini Union Hotel\\
&Flights&800&6&4800&Estimated\\
&Taxi&20&12&240&\\
&Incidentals&100&6&600&\\
&Continental breakfast&8.5&24&204&University Catering\\
&Coffee Service&32.5&4&130&\\
&Lunch&15&24&360&University Catering\\
&Dinner&30&24&720&In Champaign-Urbana\\
        \hline
        &\textbf{Total}&&&\textbf{23606}&\\
        \hline
\end{tabularx}
\caption{All expenses will support travel for visitors and workshop 
necessities.}
\end{table}
